---
title: "Predictive Machine Learning Assignment Write Up"
output: html_document
---

Machine Learning Assignment

========================================================

# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

# Data

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


# Cleaning and Processing Data

I will need to use these libraries for this assignment.

```{r}
library(randomForest)
library(caret)
library(ggplot2)
library(lattice)
```

Now that that's out of the way, let's download the data from the links above and pull it into R.

```{r}
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
              destfile = "./PML_TRAIN.csv")
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
              destfile = "./PML_TEST.csv")
PMLtrain1 <- read.csv("PML_TRAIN.csv", na.strings=c("","NA","#DIV/0"))
PMLtest1 <- read.csv("PML_TEST.csv", na.string=c("","NA", "#DIV/0")) 
```

We now have both sets to train and test with. But upon further review of the data, there are a whole lot of N/A's. I decided to remove any columns in this data that had a lot of NAs.

```{r}
clean <- colSums(is.na(PMLtrain1)) == 0
PMLtrain2 <- PMLtrain1[,clean]
PMLtest2 <- PMLtest1[,clean]
```

We're almost there! I also noticed the first 7 columns were not helpful in analyzing this data at all. Things like date, time, name, etc that would not be necessary to keep.

```{r}
head(PMLtrain2)

PMLtrain.final <- PMLtrain2[,-(1:7)]
PMLtest.final <- PMLtest2[,-(1:7)]
```

Ok, I'm very confident in this data now. Let's move on to the modeling

# Creating the Model

First, obviously we have to divide the data into a training and test set. I went with 80% training and 20% testing.

```{r}
set.seed(703)
PMLtraining <- createDataPartition(PMLtrain.final$classe, p = 0.8, list = FALSE)

PMLtrain.set <- PMLtrain.final[PMLtraining, ]
PMLtrain.validation <- PMLtrain.final[-PMLtraining, ]

```

Now that both sets of data are divided, I'm going to start building the data thanks to the randomForest library

```{r}
PML.model <- randomForest(classe ~ ., data = PMLtrain.set)
PML.model
```

This is the best model of all time. My Estimate of error is only 0.5%!

# Cross Validating our Test Set

Now I'm going to see how the trained model works on the test group we withheld. 

```{r}
PMLvalidate <- predict(PML.model, PMLtrain.validation)
confusionMatrix(PMLvalidate,PMLtrain.validation$classe)

```

Wow the accuracy is 99.64%! Time to test it on the test PML data set we grabbed earlier.

# Applying the Model on Test Data

```{r}
PMLpredict <- predict(PML.model, PMLtest.final)
PMLpredict
```

We have our answers. Now to use the given function to write out each answer to a text file.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(PMLpredict)
```

And finally.... after submitting our answers online. We got 20/20 correct! Great work gang!
